GVN-hoist: hoisting expressions from control flow

### 2016-08-04 hoist.tex
register pressure

\subsection{Reduce register pressure}
Hoisting scalars across calls is tricky because it can increase the
number of spills.

>>>>

\subsection{Reduce register pressure}
Following example explains how code hoisting can actually reduce the register pressure.
Consider the following example where the labels prefixed with 'p' represent the position of
instruction in a basic block (names prefixed with bb).

bb0: p0: b = ...
bb1: p1: c = ...
bb2: p2: if (c) jump bb3 else jump bb4
bb3: p3: a0 = b + c
bb4: p4: a1 = b + c
bb5: p5: d = phi {a0(bb3), a1(bb4)}

If we measure distance(px, py) as total instruction count in the path from the position of px to py

old-live-range = max(distance(p0, p3) + distance(p1, p3) , distance(p0, p4) + distance(p1, p4)) + distance (p3, p5)
new-live-rance = distance(p0, p2) + distance(p1, p2) + distance (p2, p5)

If the new live-range is less than the old one it will be a good candidate
for hoist. There are cases when both the ranges will be same:
\begin{enumerate}
  - There is only one operand on the right hand side e.g., an assignment operation.
  - One of the operands is not a Instruction/Register.
\end{enumerate}

In a special case where the instruction to be hoisted has the last use of its
operands then the code hoisting will always reduce the register pressure if it
has two register operands because the gain in live-range will be in the ratio of
2:1. Based on the above formulae we can also deduce that, as long as there is
one register operand in the right hand side with its last use, code hoisting
will either decrease or maintain the register pressure.

Hoisting scalars across calls is tricky because it can increase the
number of spills.
